\chapter{Data Format Compatibility}
\label{compatible}


%% TODO: Replace this note with actual changes.

{\fbox{\parbox{5in}{
{\bf *** NOTE TO USERS ***} \newline
This section of the User's Guide needs
updating.  For example, the CoastWatch IMGMAP format is no longer supported,
and cwexport and CDAT can write NetCDF 3 and 4 files.
}}



Many of the tools automatically detect the input file format and
perform the same operation for a number of different formats, for
example CoastWatch IMGMAP files and CoastWatch HDF files. The
following table shows the file format compatibility of the various
tools: \\
\\
\begin{tabular}{|l|p{5cm}|p{5cm}|}
  \hline 

  Tool & Input formats & Output formats \\ \hline 

  cdat & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & CWHDF, Binary, Text,
  ArcGIS, PNG, GIF, JPEG, PDF, GeoTIFF \\ \hline

  cwangles & CWHDF/NC & -- \\ \hline

  cwautonav & CWHDF/NC, CWF & -- \\ \hline

  cwcomposite & CWHDF/NC, CWF, TSHDF, OPeNDAP & CWHDF \\ \hline

  cwcoverage & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & PNG \\ \hline

  cwdownload & -- & -- \\ \hline

  cwexport & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & Binary, Text, ArcGIS \\
  \hline

  cwgraphics & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & CWHDF \\ \hline

  cwimport & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & CWHDF \\ \hline

  cwinfo & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & -- \\ \hline

  cwmaster & CWHDF/NC, CWF, TSHDF & CWHDF \\ \hline

  cwmath & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & CWHDF \\ \hline

  cwnavigate & CWHDF/NC, CWF & -- \\ \hline

  cwregister & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & CWHDF \\ \hline

%%\end{tabular}
%%
%%\begin{tabular}{|l|p{5cm}|p{5cm}|}
%%  \hline
%%
%%  Tool & Input formats & Output formats \\ \hline

  cwrender & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & PNG, GIF, JPEG, PDF,
  GeoTIFF \\ \hline

  cwsample & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & Text \\ \hline

  cwstats & CWHDF/NC, CWF, TSHDF, N1B, OPeNDAP & -- \\ \hline

  cwstatus & -- & -- \\ \hline

\end{tabular}

where:

\begin{description}

  \item[CWHDF/NC] HDF 4 and netCDF 3 with CoastWatch HDF or CF
  metadata (.hdf, .nc).  CF metadata support is still experimental.

  \item[CWF] CoastWatch format (.cwf) files that contain AVHRR SST, cloud,
  channel, angle, and graphics data.  These files were produced by
  CoastWatch up to 2004.

  \item[TSHDF] SeaSpace TeraScan files that have been converted to HDF
  using tdftohdf.  Only the {\tt rectangular}, {\tt polarstereo}, {\tt
  mercator}, and {\tt emercator} projections are supported, along with
  {\tt sensor\_scan} if latitude and longitude angles have been
  included in the file.

  \item[N1B] NOAA 1b AVHRR data, format versions 1 through 5, GAC or LAC
  resolutions, 8-bit, 10-bit, and 16-bit sensor word sizes with or
  without archive header.

  \item[OPeNDAP] OPeNDAP datasets with CoastWatch HDF or CF metadata,
  map projection files only.  CF metadata support is still
  experimental.  Swath datasets and {\tt RasterPixelIsPoint} mode are
  not supported.

  \item[Binary] Raw binary format.  See individual manual pages for a
  description of the binary file contents.

  \item[Text] Standard ASCII text format.  See individual manual pages
  for a description of the text file contents.

  \item[ArcGIS] ESRI ArcGIS binary grid format (32-bit IEEE floats) with
  accompanying header file.

  \item[PNG] Portable Network Graphics with optional world file.

  \item[GIF] Graphics Interchange Format with optional world file.

  \item[JPEG] Joint Photographic Experts Group with optional world file.

  \item[PDF] Adobe Portable Document Format

  \item[GeoTIFF] Tagged Image File Format with georeferencing tags.

\end{description}
