\chapter{Introduction}

The main goal of the CoastWatch Utilities software is to aid data
users in working with NOAA/NESDIS CoastWatch satellite data. 
CoastWatch data is distributed as individual files in a scientific
data format that is not recognized by standard image viewers, and the
CoastWatch Utilities are useful for manipulating data and creating images
from CoastWatch data for both recreational and scientific
applications.  CoastWatch data files contain:
\begin{enumerate}

  \item Global file attributes that describe the date/time and
  location of the earth data in the file, as well as any relevant data
  processing details.

  \item Earth data as a set of two-dimensional numerical arrays, each
  with a unique name.  These {\em variables} hold the actual
  scientific data and accompanying attributes, such as scaling factor
  and units, that describe how to use the data values.

\end{enumerate}
The CoastWatch Utilities allow users to selectively access and extract
this information in a number of ways.

\section{A Brief History}

The CoastWatch Utilities have been evolving since 1998.  The original
software was capable of working with CoastWatch IMGMAP format data
files, the standard data distribution format (limited to NOAA AVHRR
data) for CoastWatch satellite data until 2003.  The current utilities
are capable of reading CoastWatch IMGMAP, NOAA 1b AVHRR, and
CoastWatch HDF, a new format based on the NCSA Hierarchical Data
Format (HDF) that has been adopted by CoastWatch for distributing
earth data from a host of different satellites and sensors.  Following
is a sketch of the software evolution:
\begin{itemize}

  \item Version 1: 1997.  The first version created CoastWatch IMGMAP
  files from TeraScan Data Format (TDF), a proprietary format from
  SeaSpace Corporation.  It was never publicly released.

  \item Version 2: 1998-2000.  The second version was released to the
  CoastWatch data user community and worked with CoastWatch IMGMAP
  (CWF) files only.  Users could convert CWF files to other formats,
  create GIF images, perform land and cloud masking, create data
  composites, and other related tasks.  

  \item Version 3: 2001-present.  The third version was designed to have
  a more flexible data handling capability, with support for both
  CoastWatch IMGMAP and the new CoastWatch HDF format.  The CoastWatch
  Data Analysis Tool (CDAT) was integrated into the package.
  Additional capabilities were added for NOAA 1b AVHRR, level 2 swath
  style data, automatic navigational correction, ESRI shapefiles, and
  other improvements.

\end{itemize}

\section{Installation Notes}

\subsection{System Requirements}

\begin{description}

  \item[Operating system] The CoastWatch Utilities are currently
  available for Windows, Linux, Solaris, Mac OS X, and AIX only.

  \item[Java version] As of version 3.2.2, the correct Java Runtime
  Environment (JRE) is packaged with the Windows, Linux, Solaris,
  and AIX versions of the CoastWatch Utilities.  There is no need
  to install it separately -- it will be placed in a subdirectory
  of the installation directory and will only be used by the
  CoastWatch Utilities.

  For Mac OS X a JRE version {\bf 1.5.0 or higher} is required on
  the machine.  Generally an up-to-date version of Mac OS X has
  the correct Java version.  To check the current Java version,
  go to {\gui Applications} -$>$ {\gui Utilities} in the Finder and
  start the Terminal.  Type \verb+java -version+, hit enter, and
  look at the first line of output.  If the version does not
  start with 1.5.0 or greater, run {\gui Software Update} from the
  \macsym{apple} menu or visit \url{http://www.apple.com/java} to
  download manually.

  \item[Disk space] A minimum of 150 Mb is required for the installed
  software. We recommended at least 300 Mb of space in total to
  allow for downloading and manipulating satellite datasets. More
  disk space may be required for larger datasets.

  \item[Memory] A minimum of 128 Mb is required, while 256 Mb is
  recommended. More memory may be required depending on the number of
  concurrent processes running on the computer.

\end{description}

\subsection{Downloading}

To download, visit the CoastWatch central web site: 
\begin{quote}
  \url{http://coastwatch.noaa.gov/cw\_cwfv3.html}
\end{quote}
and download the package appropriate for your operating system. To
install, read one of the following sections depending on your
platform.

\subsection{Installing on Windows}

If upgrading to a new version, there is no need to uninstall the
previous version -- the new installation setup program will handle
the details.  To install the new version, simply double-click the
downloaded installation package. The setup program will install the
software in a user-specified directory, and add entries to the Start
Menu.  The User's Guide (this document) is also added to the Start
Menu.

If you use the command line tools you may want to add the installed
executable directory to the path for easier command line execution.
On Windows XP, go to the Windows {\gui Start Menu} and click {\gui
Control Panel} -$>$ {\gui Performance and Maintenance} -$>$ {\gui
System}, select the {\gui Advanced} tab and click the {\gui Environment
Variables} button. Edit the entry for the {\gui Path} variable under
{\gui System variables} and add the full path to the installed
executable directory, for example {\tt C:$\backslash$Program
Files$\backslash$CoastWatch Utilities$\backslash$bin}. Other Windows
versions have a similar procedure for adding a path.

\subsection{Installing on Mac OS X}

The Mac OS X installation package is a disk image (DMG) file that
contains support for both PowerPC and Intel processors.  
Download the DMG file, double-click to mount it, and then
double-click the installer program.  When upgrading, the new version
of the utilities will automatically uninstall the old version.

If you use the command line tools you may want to add the installed
executable directory to your path for easier command line execution.
In a Terminal session, add the following lines to the {\tt .profile}
file:
\begin{quote}
  {\tt export PATH=\$\{PATH\}:"/Applications/CoastWatch Utilities/bin"}
\end{quote}
and then open a new Terminal window for the changes to take effect.

\subsection{Installing on Unix}

For the Linux RPM file, run the rpm command as root:
\begin{quote}
  {\file rpm -Uvh} FILENAME.RPM
\end{quote}
which will install or upgrade the software into {\tt /opt/cwf} and create
links to the executables in {\tt /usr/local/bin}. For 
the self-installing shell script, you can
install as a normal user or as root by executing the shell script
and following the instructions in the graphical setup wizard.
For the tar archive file (non-graphical install),
extract the archive as follows:
\begin{quote}
  {\file gunzip} FILENAME.TAR.GZ \\
  {\file tar -xf} FILENAME.TAR
\end{quote}
You may also want to add the installed executable directory to
your {\tt \$PATH} environment variable for easier command line
execution, for example:
\begin{quote}
  {\tt setenv PATH \$\{PATH\}:/opt/cwf/bin}~(in {\tt .cshrc}) \\
  {\tt export PATH=\$\{PATH\}:/opt/cwf/bin}~(in {\tt .bashrc})
\end{quote}

\section{Using the Software}

The CoastWatch Utilities are made up of various individual tools.
Graphical tools allow you to point and click, working with data
interactively; the graphical tools have a built-in help system
with brief details on each key feature.  To complement these,
there are also command line tools that allow you to process data
using a text-only command prompt or scripting language.  Call any
command line tool with the \longoption{help} option to get a
short synopsis of parameters.  See the manual pages of
\autoref{manual} for a complete discussion on the command line
parameters for both graphical and command line tools.

Functionally, the CoastWatch Utilities are designed to meet the data
processing and rendering needs of many different types of data users.
The individual tools have been developed from both in-house
requirements and data user suggestions.  The functionality of the
tools may be divided into several categories based on data processing
task:
\begin{description}

  \item[Information and Statistics] File contents, statistics
  computations on variables (for example min, max, mean, standard
  deviation), direct access to raw file and variable attributes.

  \item[Data Processing] Data format conversions,
  compositing, generic variable math, data sampling.

  \item[Graphics and Visualization] Interactive
  visualization/analysis, batch image rendering, ancillary graphics
  creation such as data coverage maps, grids, coastlines, landmasks.
    
  \item[Registration and Navigation] Resampling of data from one
  projection to another, interactive generation of region masters,
  manual and automatic navigational correction, computation of solar
  and earth location angles.

  \item[Network] Data download and server status.

\end{description}

\section{Third Party Software}
\label{third}

There are a number of other software packages than can
be used to read data from CoastWatch HDF data files.  They are {\em
generic} in the sense that they can read the numerical and text data,
but they cannot interpret the metadata conventions used by CoastWatch.
They are well suited to advanced users who want to have more detail on
HDF data file contents.  You should always refer to the CoastWatch
HDF metadata specification of \autoref{metadata} when using third
party software with CoastWatch HDF files.

The HDF Group (THG), originally part of the National Center for
Supercomputing Applications (NCSA), is the main source of information
on the HDF format specification and software libraries.  You can
download tools for working with HDF from the main web site:
\begin{quote}
  \url{http://www.hdfgroup.org}
\end{quote} 
Many of the tools are command line, but HDFView is a useful graphical
tool.

Other non-NCSA software has been HDF-enabled by linking against the
NCSA HDF libraries.  Among those packages are IDL, Matlab, and
OPeNDAP.  A more complete list of non-NCSA software may be found at:
\begin{quote}
  \url{http://www.hdfgroup.org/tools.html}
\end{quote}

% TODO: Need to add something here about where to get CoastWatch data.


